---
title: "Run experiment"
author: "Francis Durnin-Vermette"
date: "2023/2/10"
output:
  html_notebook:
    toc: true
    number_sections: true
    df_print: paged
bibliography: bib.bib
editor_options:
  chunk_output_type: inline
---
<!-- Setup chunk: -->
```{r include=FALSE}
library(ggplot2)
library(knitr)
library(dplyr)
library(geojson)
library(tibble)
library(purrr)
library(stringr)
library(here)
library(lubridate)
library(tidyr)
dir.create(tempdir()) #This fixes a bug if the temporary directory is not found
here::i_am("walkthrough_icbm.rmd")
source(here::here("r/icbm_calculate_re.r"))
source(here::here("r/icbm_run.r"))
source(here::here("r/read_climate_data.r"))
source(here::here("r/spinup.r"))
source(here::here("r/bayesian_gsa.r"))
source(here::here("r/bayesian_sir.r"))
source(here::here("r/bayesian_validation.r"))
source(here::here("r/bayesian_uncertainty.r"))
```

Assign experimental parameters:
```{r}
experiment_driver <- "location_name"
experiment_driver_values <- c("Ellerslie", "Harrow")
experiment_model <- "ipcct2"
experiment_time <- Sys.time() %>%
	format("%Y%m%dT%H%M%S")
```

Choose sample sizes for the analysis:
```{r}
calibration_sample_percentage <- 10
gsa_sample_size <- 3
sir_sample_size <- 4
sir_resample_size <- 3
```

Create folders needed for analysis:
```{r}
experiment_driver_sanitized <- gsub("[^A-Za-z0-9 ]","", experiment_driver)
experiment_driver_values_sanitized <- gsub("[^A-Za-z0-9 ]","",experiment_driver_values)

experiment_name <- paste0(c(experiment_driver_sanitized, experiment_driver_values_sanitized), collapse = "-")

experiment_folder <- paste0(experiment_name, "_", experiment_time, "_", "gsa", gsa_sample_size, "_", "sir", sir_sample_size, "-", sir_resample_size, "_", "cal", calibration_sample_percentage)
experiment_path <- here("results", experiment_model, experiment_folder)

dir.create(here(experiment_path))
dir.create(here(experiment_path, "inputs"))
dir.create(here(experiment_path, "gsa"))
dir.create(here(experiment_path, "sir"))
dir.create(here(experiment_path, "validation"))
dir.create(here(experiment_path, "uncertainty"))
```


# Input data

## Get site data:
```{r}
# Read SLC polygon IDs
polyids_all <- readr::read_csv(here("data","climate_data","Documentation", "SLC_PCPTREGION.csv"))
# Read site data
site_data <- readr::read_csv(here("data","LTE_Master_beta_mar15_addedcolumns.csv")) %>%
	left_join(polyids_all, by=c("latitude" = "Lat", "longitude" = "Long")) %>%
	mutate(tillage = ifelse(is.na(tillage), "CT", tillage))

# Save to file
readr::write_csv(site_data, file=here(experiment_path,"inputs", "site_data.csv"))
```
## Add specific site values from lookup tables
e.g. cinput_lignin_px

## Calculate the SOC at 30cm depth
```{r}
site_data_30cm <- site_data %>%
	mutate(soc_tha_30cm = soc_tha)
```


## Remove sites where there are no measured values
```{r}
site_data_filtered <- site_data_30cm %>%
	group_by(TrtID, replication_number) %>%
	filter(!all(is.na(soc_tha_30cm)))

```


## Only select data associated with the driver:
```{r}
site_data_driver <- site_data_filtered %>%
	filter(.data[[experiment_driver]] %in% experiment_driver_values)

# Save to file
site_data_driver %>%
	readr::write_csv(file=here(experiment_path,"inputs","site_data_driver.csv"))
```


## Split the data into calibration and validation sets:
```{r}
site_data_driver_n <- site_data_driver %>%
	group_by(TrtID, replication_number) %>%
	n_groups

calibration_sample_indices <- sample.int(site_data_driver_n, ceiling(site_data_driver_n * calibration_sample_percentage/100))
evaluation_sample_indices <- setdiff(1:site_data_driver_n, calibration_sample_indices)


calibration_data <- site_data_driver %>%
	group_by(TrtID, replication_number) %>%
	group_split %>%
	`[`(calibration_sample_indices)

evaluation_data <- site_data_driver %>%
	group_by(TrtID, replication_number) %>%
	group_split %>%
	`[`(evaluation_sample_indices)

# Save to file
calibration_data %>%
	bind_rows() %>%
	readr::write_csv(file=here(experiment_path,"inputs","site_data_calibration.csv"))
evaluation_data %>%
	bind_rows() %>%
	readr::write_csv(file=here(experiment_path,"inputs","site_data_evaluation.csv"))
```

## Get climate data:
```{r}
climate_data <- read_climate_data(climate_data_directory = here("data","climate_data","W9param_TablesCleaned"))

# Save to file
readr::write_csv(climate_data, file=here(experiment_path,"inputs","climate_data.csv"))
```

## Parameter ranges used in sensitivity analysis and bayesian calibration:
```{r}
param_bounds <- read.csv(here("data","ipcct2_parameters_gsa.csv"), stringsAsFactors = FALSE)

# Save to file
readr::write_csv(param_bounds, file=here(experiment_path,"inputs","parameter_ranges.csv"))
```

## Run spinup:
```{r}
# Calibration
calibration_data_spinup <- calibration_data %>%
	map(~slice(., 1:10))

# Get initial C
initial_c_calibration_measured <- calibration_data_spinup %>%
	map(~.$initial_soc_tha[1]*10) %>%
	map(~ifelse(is.na(.), 1000, .))

# Run spinup
initial_c_calibration_spinup <- map2(.x = calibration_data_spinup,
																		 .y = initial_c_calibration_measured,
																		 ~spinup(site_data = .x, 
																		 				climate_data = climate_data,
																		 				initial_c = .y,
																		 				model = "ipcct2"))

# Evaluation
evaluation_data_spinup <- evaluation_data %>%
	map(~slice(., 1:10))

# Get initial C
initial_c_evaluation_measured <- evaluation_data %>%
	map(~.$initial_soc_tha[1]*10) %>%
	map(~ifelse(is.na(.), 1000, .))

# Run spinup
initial_c_evaluation_spinup <- map2(.x = evaluation_data_spinup,
																		 .y = initial_c_evaluation_measured,
																		 ~spinup(site_data = .x, 
																		 				climate_data = climate_data,
																		 				initial_c = .y,
																		 				model = "ipcct2"))

```

# Sensitivity Analysis
```{r}
# Sobol-Jansen'
sensitivity_calibration_sj <- gsa(site_data = calibration_data,
																	climate_data = climate_data,
																	initial_c = initial_c_calibration_spinup,
																	parameter_bounds = param_bounds,
																	sample_size = gsa_sample_size,
																	method = "soboljansen")
# Save to file
sensitivity_calibration_sj %>%
	mutate(driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 gsa_sample_n = gsa_sample_size) %>%
	readr::write_csv(file=here(experiment_path,"gsa","sj.csv"))

# Fast99
sensitivity_calibration_f99 <- gsa(site_data = calibration_data,
																	 climate_data = climate_data,
																	 initial_c = initial_c_calibration_spinup,
																	 parameter_bounds = param_bounds,
																	 sample_size = gsa_sample_size,
																	 method = "fast99")
# Save to file
sensitivity_calibration_f99 %>%
	mutate(driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 gsa_sample_n = gsa_sample_size) %>%
	readr::write_csv(file=here(experiment_path,"gsa","f99.csv"))
```

## Graphing:
```{r}
# Template
sensitivity_graph_template <- ggplot(data=NULL) + 
	coord_flip() +
	theme_bw()

# Sobol-Jansen
# First-order sensitivity
sensitivity_graph_first_sj <- sensitivity_graph_template %+% sensitivity_calibration_sj +
	aes(x = reorder(params, singsi),
			y = singsi,
			ymax=singsi_uci,
			ymin=singsi_lci) +
	geom_bar(stat='identity', fill="grey", alpha=0.70 ) +
	geom_errorbar(width=0.2, linewidth=1, color="black") +
	ylab("First-Order Sensitivity Index") +
	xlab("Parameters")


# Total order sensitivity
sensitivity_graph_total_sj <- sensitivity_graph_template %+% sensitivity_calibration_sj +
	aes(x = reorder(params, totsi),
			y = totsi,
			ymax=totsi_uci,
			ymin=totsi_lci) +
	geom_bar(stat='identity', fill="grey", alpha=0.70 ) +
	geom_errorbar(width=0.2, linewidth=1, color="black") +
	ylab("Total-Order Sensitivity Index") +
	xlab("Parameters")

# Both
sensitivity_calibration_sj_pivoted <- sensitivity_calibration_sj%>%
	pivot_longer(c("totsi", "singsi"), names_to = "order", values_to = "index")

sensitivity_graph_both_sj <-  sensitivity_graph_template %+% sensitivity_calibration_sj_pivoted +
	aes(x = reorder(params, order),
			y = index) +
	geom_bar(aes(fill = order), stat='identity', position="dodge", alpha=0.7) +
	geom_errorbar(aes(ymax=singsi_uci, ymin=singsi_lci), width=0.2, linewidth=1, color="black") +
	geom_errorbar(aes(ymax=totsi_uci, ymin=totsi_lci), width=0.2, linewidth=1, color="#888888") +
	ylab("Sensitivity Index") +
	xlab("Parameters")

# Fast99
# Main
sensitivity_graph_main_f99 <- sensitivity_graph_template %+% sensitivity_calibration_f99 +
	aes(x = reorder(params, main),
			y = main) + 
	xlab("Parameters") +
	ylab("Main Effect Sensitivity Index") +
	geom_bar(stat='identity', fill="grey", alpha=0.70)
# Interactions
sensitivity_graph_interaction_f99 <- sensitivity_graph_template %+% sensitivity_calibration_f99 +
	aes(x = reorder(params, interactions),
			y = interactions) + 
	xlab("Parameters") +
	ylab("Interaction Effect Sensitivity Index") +
	geom_bar(stat='identity', fill="grey", alpha=0.70)
# Both
sensitivity_calibration_f99_pivoted <- sensitivity_calibration_f99 %>%
	pivot_longer(c("main", "interactions"), names_to = "order", values_to = "index")

sensitivity_graph_both_f99 <-  sensitivity_graph_template %+% sensitivity_calibration_f99_pivoted +
	aes(x = reorder(params, index),
			y = index) +
	geom_bar(aes(fill = order), stat='identity', position="dodge", alpha=0.7) +
	ylab("Sensitivity Index") +
	xlab("Parameters")

# Save to file
ggsave(plot = sensitivity_graph_first_sj, filename = here(experiment_path,"gsa","sj_first.png"))
ggsave(plot = sensitivity_graph_total_sj, filename = here(experiment_path,"gsa","sj_total.png"))
ggsave(plot = sensitivity_graph_both_sj, filename = here(experiment_path,"gsa","sj_both.png"))

ggsave(plot = sensitivity_graph_main_f99, filename = here(experiment_path,"gsa","f99_main.png"))
ggsave(plot = sensitivity_graph_interaction_f99, filename = here(experiment_path,"gsa","f99_interaction.png"))
ggsave(plot = sensitivity_graph_both_f99, filename = here(experiment_path,"gsa","f99_both.png"))
```

# Sampling Importance Resampling (SIR)
```{r}
sir_calibration <- sir(site_data = calibration_data,
											 climate_data = climate_data,
											 initial_c = initial_c_calibration_spinup,
											 parameter_bounds = param_bounds,
											 sample_size = sir_sample_size,
											 resample_size = sir_resample_size)

readr::write_csv(sir_calibration$prior, file=here(experiment_path, "sir", "prior.csv"))
readr::write_csv(sir_calibration$posterior, file=here(experiment_path, "sir", "posterior.csv"))

# Find maximum a posteriori (MAP)
get_map <- function(x) {
	dens <- x %>%
		density
	return(dens$x[which.max(dens$y)])
}

sir_table <- sir_calibration$posterior %>%
	pivot_longer(cols = !SampleID) %>%
	group_by(parameter = name) %>%
	summarise(`2.5%` = quantile(value, probs = 0.025),
						`25%` = quantile(value, probs = 0.25),
						median = quantile(value, probs = 0.50),
						`75%` = quantile(value, probs = 0.75),
						`97.5%` = quantile(value, probs = 0.975),
						map = get_map(value),
						mean = mean(value))

sir_table %>%
	mutate(driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 sir_sample_n = sir_sample_size,
				 sir_resample_n = sir_resample_size) %>%
	readr::write_csv(file= here(experiment_path,"sir","summary.csv"))
```

## Graphing:
```{r}
prior <- sir_calibration$prior %>%
	pivot_longer(cols = !`SampleID`)
posterior <- sir_calibration$posterior %>%
	pivot_longer(cols = !`SampleID`)

sir_graph <- ggplot() +
	geom_density(data = posterior, aes(value), col = NA, fill = "red", alpha = 0.2) +
	geom_density(data = prior, aes(value), col = NA, fill = "blue", alpha = 0.2) +
	facet_wrap(~name, scales = "free", ncol = 3) +
	theme_bw() +
	theme(panel.border = element_rect(colour = "black", fill = NA), panel.grid.major = element_blank(),
				axis.title.x=element_blank(),panel.grid.minor = element_blank()) +
	scale_y_continuous(expand = c(0,0)) + scale_x_continuous(expand = c(0,0))

ggsave(plot = sir_graph, filename = here(experiment_path,"sir","sir.png"))
```

# Validation

## Make model returns for evaluation data set
```{r}
modelled_evaluation_prior <- run_loglike_parallel(site_data = evaluation_data,
																									climate_data = climate_data,
																									initial_c = initial_c_evaluation_spinup,
																									parameter_sample = sir_calibration$prior) %>%
	`[[`("model_return")

modelled_evaluation_posterior <- run_loglike_parallel(site_data = evaluation_data,
																											climate_data = climate_data,
																											initial_c = initial_c_evaluation_spinup,
																											parameter_sample = sir_calibration$posterior) %>%
	`[[`("model_return")
```

## Save model runs for each data set
```{r}
# Evaluation
modelled_evaluation_prior %>%
	bind_rows %>%
	readr::write_csv(file=here(experiment_path, "validation", "modelled_evaluation_prior.csv"))
modelled_evaluation_posterior %>%
	bind_rows %>%
	readr::write_csv(file=here(experiment_path, "validation", "modelled_evaluation_prior.csv"))

# Calibration
sir_calibration$model_return_prior %>%
	bind_rows %>%
	readr::write_csv(file=here(experiment_path, "validation", "modelled_calibration_prior.csv"))
sir_calibration$model_return_posterior %>%
	bind_rows %>%
	readr::write_csv(file=here(experiment_path, "validation", "modelled_calibration_posterior.csv"))
```


## Graphing SOC and ΔSOC
```{r}
# Generic function for building graphs
make_soc_graphs <- function(model_return_data) 
{
	soc_graph <- ggplot(model_return_data, aes(x = year)) +
		geom_point(aes(y = soc_total)) +
		geom_point(aes(y = soc_tha_30cm), color="red") +
		facet_wrap(vars(treatment_name)) +
		ylab("SOC (t/ha)") +
		theme_bw()
	
	soc_delta_cumulative_graph <- ggplot(model_return_data, aes(x = year)) +
		geom_point(aes(y = soc_total_delta_cumulative)) +
		geom_point(aes(y = soc_tha_30cm_delta_cumulative), color="red") +
		facet_wrap(vars(treatment_name)) +
		ylab("ΔSOC (t/ha)") +
		theme_bw()
	
	soc_delta_progressive_graph <- ggplot(model_return_data, aes(x = year)) +
		geom_point(aes(y = soc_total_delta_progressive)) +
		geom_point(aes(y = soc_tha_30cm_delta_progressive), color="red") +
		facet_wrap(vars(treatment_name)) +
		ylab("Year-by-year ΔSOC (t/ha)") +
		theme_bw()
	
	data_name <- deparse(substitute(model_return_data))
	
	ggsave(plot = soc_graph, filename = here(experiment_path,"validation",paste0(data_name,"_soc.png")))
	ggsave(plot = soc_delta_cumulative_graph, filename = here(experiment_path,"validation",paste0(data_name,"_soc_delta_cumulative.png")))
	ggsave(plot = soc_delta_progressive_graph, filename = here(experiment_path,"validation",paste0(data_name,"_soc_delta_progressive.png")))
	
	return(list(
		soc_graph = soc_graph,
		soc_delta_cumulative_graph = soc_delta_cumulative_graph,
		soc_delta_progressive_graph = soc_delta_progressive_graph
	))
}

```


```{r}
# Conduct validation on all data sets
validation_evaluation_prior <- modelled_evaluation_prior %>%
	purrr::map(validation)
validation_evaluation_posterior <- modelled_evaluation_posterior %>%
	purrr::map(validation)
validation_calibration_prior <- sir_calibration$model_return_prior %>%
	purrr::map(validation)
validation_calibration_posterior <- sir_calibration$model_return_posterior %>%
	purrr::map(validation)

# Get soc and delta soc values
stocks_evaluation_prior <- validation_evaluation_prior %>%
	purrr::map(~.[["stocks"]]) %>%
	bind_rows
stocks_evaluation_posterior <- validation_evaluation_posterior %>%
	purrr::map(~.[["stocks"]]) %>%
	bind_rows
stocks_calibration_prior <- validation_calibration_prior %>%
	purrr::map(~.[["stocks"]]) %>%
	bind_rows
stocks_calibration_posterior <- validation_calibration_posterior %>%
	purrr::map(~.[["stocks"]]) %>%
	bind_rows

# Make graphs
soc_graphs_evaluation_prior <- make_soc_graphs(stocks_evaluation_prior)
soc_graphs_evaluation_posterior <- make_soc_graphs(stocks_evaluation_posterior)
soc_graphs_calibration_prior <- make_soc_graphs(stocks_calibration_prior)
soc_graphs_calibration_posterior <- make_soc_graphs(stocks_calibration_posterior)
```

## Calculating Validation statistics
```{r}
validationstats_evaluation_prior <- validation_evaluation_prior %>%
	purrr::map(~.[["validation"]]) %>%
	bind_rows %>%
	mutate(distribution = "evaluation_prior",
				 driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 gsa_sample_n = gsa_sample_size,
				 sir_sample_n = sir_sample_size,
				 sir_resample_n = sir_resample_size)

validationstats_evaluation_posterior <- validation_evaluation_posterior %>%
	purrr::map(~.[["validation"]]) %>%
	bind_rows %>%
	mutate(distribution = "evaluation_posterior",
				 driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 gsa_sample_n = gsa_sample_size,
				 sir_sample_n = sir_sample_size,
				 sir_resample_n = sir_resample_size)

validationstats_calibration_prior <- validation_calibration_prior %>%
	purrr::map(~.[["validation"]]) %>%
	bind_rows %>%
	mutate(distribution = "calibration_prior",
				 driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 gsa_sample_n = gsa_sample_size,
				 sir_sample_n = sir_sample_size,
				 sir_resample_n = sir_resample_size)

validationstats_calibration_posterior <- validation_calibration_posterior %>%
	purrr::map(~.[["validation"]]) %>%
	bind_rows %>%
	mutate(distribution = "calibration_posterior",
				 driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 gsa_sample_n = gsa_sample_size,
				 sir_sample_n = sir_sample_size,
				 sir_resample_n = sir_resample_size)

# Save to file
readr::write_csv(validationstats_evaluation_prior, file = here(experiment_path,"validation","validation_evaluation_prior.csv"))
readr::write_csv(validationstats_evaluation_posterior, file = here(experiment_path,"validation","validation_evaluation_posterior.csv"))
readr::write_csv(validationstats_calibration_prior, file = here(experiment_path,"validation","validation_calibration_prior.csv"))
readr::write_csv(validationstats_calibration_posterior, file = here(experiment_path,"validation","validation_calibration_posterior.csv"))
```

## Graphing validation statistics
```{r}
# Evaluation
validation_evaluation <- bind_rows(validationstats_evaluation_prior, validationstats_evaluation_posterior)

validation_evaluation_violin_graph <- ggplot(validation_evaluation, aes(y=rmse_stocks, x=distribution)) +
	geom_violin() +
	theme_bw()
ggsave(plot = validation_evaluation_violin_graph, filename = here(experiment_path,"validation","validation_evaluation.png"))

# Calibration
validation_calibration <- bind_rows(validationstats_calibration_prior, validationstats_calibration_posterior)

validation_calibration_violin_graph <- ggplot(validation_calibration, aes(y=rmse_stocks, x=distribution)) +
	geom_violin() +
	theme_bw()
ggsave(plot = validation_calibration_violin_graph, filename = here(experiment_path,"validation","validation_calibration.png"))
```


# Uncertainty Analysis
```{r}
uncertainty_calibration_prior <- montecarlo(site_data = calibration_data,
																						climate_data = climate_data,
																						initial_c = initial_c_calibration_spinup,
																						distribution = sir_calibration$prior)
uncertainty_calibration_posterior <- montecarlo(site_data = calibration_data,
																								climate_data = climate_data,
																								initial_c = initial_c_calibration_spinup,
																								distribution = sir_calibration$posterior)

# Write to file
# Prior
readr::write_csv(uncertainty_calibration_prior$median, file=here(experiment_path,"uncertainty", "median_prior.csv"))
readr::write_csv(uncertainty_calibration_prior$montecarlo, file=here(experiment_path,"uncertainty", "montecarlo_prior.csv"))
# Posterior
readr::write_csv(uncertainty_calibration_posterior$median, file=here(experiment_path,"uncertainty", "median_posterior.csv"))
readr::write_csv(uncertainty_calibration_posterior$montecarlo, file=here(experiment_path,"uncertainty", "montecarlo_posterior.csv"))
```

## Graphing
```{r}
# 1:1 simulated:measured graph
uncertainty_graph <- ggplot(data = NULL, aes(x=soc_total/10, y=soc_tha*(30/(soil_depth_max_cm - soil_depth_min_cm)))) +
	geom_point(data=uncertainty_calibration_prior$montecarlo, shape=4, color = "lightgrey") +
	geom_point(data=uncertainty_calibration_posterior$montecarlo, shape=3, color = "darkgrey") +
	geom_point(data=uncertainty_calibration_posterior$median, color="black") +
	geom_abline(intercept = 0, slope = 1) +
	coord_fixed(ratio = 1, xlim = c(0,400), ylim = c(0,400), expand = TRUE, clip="off") +
	theme_bw()

ggsave(plot = uncertainty_graph, filename = here(experiment_path,"uncertainty","uncertainty.png"))
```

```{r}
ribbon_points_posterior <- uncertainty_calibration_posterior$montecarlo %>%
	group_by(year, treatment_name) %>%
	select(year, treatment_name, soc_total) %>%
	summarise(soc_total_max = max(soc_total),
						soc_total_min = min(soc_total))

ribbon_points_prior <- uncertainty_calibration_prior$montecarlo %>%
	group_by(year, treatment_name) %>%
	select(year, treatment_name, soc_total) %>%
	summarise(soc_total_max = max(soc_total),
						soc_total_min = min(soc_total))

ggplot(data = NULL, aes(x=year)) +
	geom_ribbon(data=ribbon_points_prior, aes(ymax=soc_total_max/10, ymin=soc_total_min/10), fill="lightgrey") +
	geom_ribbon(data=ribbon_points_posterior, aes(ymax=soc_total_max/10, ymin=soc_total_min/10), fill="darkgrey") +
	geom_point(data=uncertainty_calibration_posterior$montecarlo, aes(y=soc_tha*(30/(soil_depth_max_cm - soil_depth_min_cm))), color = "black") +
	geom_line(data=uncertainty_calibration_posterior$median, aes(y=soc_total/10), color = "black") +
	facet_wrap(vars(treatment_name)) +
	theme_bw()

```


Save the workspace image:
```{r}
save.image(file = here(experiment_path,"finished_workspace.RData"))
```



