---
title: "Run experiment"
author: "Francis Durnin-Vermette"
date: "2023/2/10"
output:
  html_notebook:
    toc: true
    number_sections: true
    df_print: paged
bibliography: bib.bib
editor_options:
  chunk_output_type: inline
---
<!-- Setup chunk: -->
```{r include=FALSE}
library(ggplot2)
library(knitr)
library(dplyr)
library(geojson)
library(tibble)
library(purrr)
library(stringr)
library(here)
library(lubridate)
library(tidyr)
dir.create(tempdir()) #This fixes a bug if the temporary directory is not found
here::i_am("walkthrough_icbm.rmd")
source(here::here("r/icbm_calculate_re.r"))
source(here::here("r/icbm_run.r"))
source(here::here("r/read_climate_data.r"))
source(here::here("r/spinup.r"))
source(here::here("r/bayesian_gsa.r"))
source(here::here("r/bayesian_sir.r"))
source(here::here("r/bayesian_validation.r"))
source(here::here("r/bayesian_uncertainty.r"))
```

Assign experimental parameters:
```{r}
experiment_name <- "test"
experiment_model <- "ipcct2"
experiment_time <- Sys.time() %>%
	format("%Y%m%dT%H%M%S")
```

Create folders needed for analysis:
```{r}
experiment_folder <- paste(experiment_name, experiment_time,  sep="_")
experiment_path <- here("results", experiment_model, experiment_folder)

dir.create(here(experiment_path))
dir.create(here(experiment_path, "code"))
dir.create(here(experiment_path, "figures"))
dir.create(here(experiment_path, "inputs"))
dir.create(here(experiment_path, "outputs"))
```


# Input data

## Get site data:
```{r}
# Read SLC polygon IDs
polyids_all <- readr::read_csv(here("data","climate_data","Documentation", "SLC_PCPTREGION.csv"))
# Read site data
site_data <- readr::read_csv(here("data","all_experiments_dummy.csv")) %>%
	#site_data <- readr::read_csv(here("data/lte_master_beta_dec02.csv")) %>%
	left_join(polyids_all, by=c("latitude" = "Lat", "longitude" = "Long")) %>%
	mutate(tillage = ifelse(is.na(tillage), "CT", tillage))

# Save to file
readr::write_csv(site_data, file=paste0(experiment_path,"/inputs/site_data.csv"))
```


## Split the data into calibration and validation sets:
```{r}
calibration_data <- site_data %>%
	filter(location_name %in% c("Ellerslie", "Harrow")) %>%
	group_by(location_name, treatment_name, replication_number) %>%
	group_split
calibration_data <- calibration_data[1:5]
validation_data <- site_data %>%
	filter(!location_name %in% c("Ellerslie", "Harrow")) %>%
	group_by(location_name, treatment_name, replication_number) %>%
	group_split

# Save to file
calibration_data %>%
	bind_rows() %>%
	readr::write_csv(file=here(experiment_path,"inputs","site_data_calibration.csv"))
validation_data %>%
	bind_rows() %>%
	readr::write_csv(file=here(experiment_path,"inputs","site_data_validation.csv"))
```

## Get climate data:
```{r}
climate_data <- read_climate_data(climate_data_directory = here("data","climate_data","W9param_TablesCleaned"))

# Save to file
readr::write_csv(climate_data, file=here(experiment_path,"inputs","climate_data.csv"))
```

## Parameter ranges that are used for sensitivity analysis and bayesian calibration:
```{r}
param_bounds <- read.csv(here("data","ipcct2_parameters_gsa.csv"), stringsAsFactors = FALSE)

# Save to file
readr::write_csv(param_bounds, file=here(experiment_path,"inputs","parameter_ranges.csv"))
```

## Run spinup:
```{r}
calibration_data_spinup <- calibration_data %>%
	map(~slice(., 1:10))

# Get initial C
initial_c_calibration_measured <- calibration_data %>%
	map(~.$initial_soc_tha[1]*10) %>%
	map(~ifelse(is.na(.), 1000, .))

# Run spinup
initial_c_calibration_spinup <- map2(.x = calibration_data_spinup,
																		 .y = initial_c_calibration_measured,
																		 ~spinup(site_data = .x, 
																		 				climate_data = climate_data,
																		 				initial_c = .y,
																		 				model = "ipcct2"))

```

# Sensitivity Analysis
```{r}
sensitivity_calibration <- gsa(site_data = calibration_data,
															 climate_data = climate_data,
															 initial_c = initial_c_calibration_spinup,
															 parameter_bounds = param_bounds,
															 sample_size = 3,
															 method = "soboljansen")

# Save to file
readr::write_csv(sensitivity_calibration, file=here(experiment_path,"outputs","sensitivity.csv"))
```

## Graphing:
```{r}
# First-order sensitivity
sensitivity_graph_first <- ggplot(sensitivity_calibration, aes(x = reorder(params, -singsi), y = singsi, ymax=singsi.uci, ymin=singsi.lci)) + 
	xlab("Parameters") +
	ylab("First-Order Sensitivity Index") +
	geom_errorbar(width=0.2, linewidth=1, color="black") +
	geom_bar(stat='identity', fill="grey", alpha=0.70 ) +
	coord_flip() +
	theme_bw()

# Total order sensitivity
sensitivity_graph_total <- ggplot(sensitivity_calibration, aes(x = reorder(params, -singsi), y = totsi, ymax=totsi.uci, ymin=totsi.lci))+ 
	xlab("Parameters") +
	ylab("Total-Order Sensitivity Index") + 
	geom_errorbar(width=0.2, linewidth=1, color="black")+
	geom_bar(stat='identity', fill="grey", alpha=0.70) +
	coord_flip() +
	theme_bw()

# Both
sensitivity_graph_both <- ggplot(sensitivity_calibration, aes(x = reorder(params, -singsi), y = totsi, ymax=totsi.uci, ymin=totsi.lci)) +
	geom_bar(aes(y = singsi), stat='identity', fill="black", alpha=0.7)+
	geom_bar(aes(y = totsi), stat='identity', fill="grey", alpha=0.7)+
	geom_errorbar(width=0.2, linewidth=1)+
	coord_flip() +
	theme_bw()

ggsave(plot = sensitivity_graph_first, filename = here(experiment_path,"figures","sensitivity_first.png"))
ggsave(plot = sensitivity_graph_total, filename = here(experiment_path,"figures","sensitivity_total.png"))
ggsave(plot = sensitivity_graph_both, filename = here(experiment_path,"figures","sensitivity_both.png"))
```

# Sampling Importance Resampling (SIR)
```{r}
sir_calibration <- sir(site_data = calibration_data,
											 climate_data = climate_data,
											 initial_c = initial_c_calibration_spinup,
											 parameter_bounds = param_bounds,
											 sample_size = 4,
											 resample_size = 3)

readr::write_csv(sir_calibration$prior, file=paste0(experiment_path,"/outputs/prior.csv"))
readr::write_csv(sir_calibration$posterior, file=paste0(experiment_path,"/outputs/posterior.csv"))

# Find maximum a posteriori (MAP)
get_map <- function(x) {
	dens <- x %>%
		density
	return(dens$x[which.max(dens$y)])
}

sir_table <- sir_calibration$posterior %>%
	pivot_longer(cols = !SampleID) %>%
	group_by(parameter = name) %>%
	summarise(`2.5%` = quantile(value, probs = 0.025),
						`25%` = quantile(value, probs = 0.25),
						median = quantile(value, probs = 0.50),
						`75%` = quantile(value, probs = 0.75),
						`97.5%` = quantile(value, probs = 0.975),
						map = get_map(value))

readr::write_csv(sir_table, file=paste0(experiment_path,"/outputs/sir_summary.csv"))
```

## Graphing:
```{r}
prior <- sir_calibration$prior %>%
	pivot_longer(cols = !`SampleID`)
posterior <- sir_calibration$posterior %>%
	pivot_longer(cols = !`SampleID`)

sir_graph <- ggplot() +
	geom_density(data = posterior, aes(value), col = NA, fill = "red", alpha = 0.2) +
	geom_density(data = prior, aes(value), col = NA, fill = "blue", alpha = 0.2) +
	facet_wrap(~name, scales = "free", ncol = 3) +
	theme_bw() +
	theme(panel.border = element_rect(colour = "black", fill = NA), panel.grid.major = element_blank(),
				axis.title.x=element_blank(),panel.grid.minor = element_blank()) +
	scale_y_continuous(expand = c(0,0)) + scale_x_continuous(expand = c(0,0))

ggsave(plot = sir_graph, filename = here(experiment_path,"figures","sir.png"))
```

# Validation

## Graphing results

## Calculating Validation statistics
```{r}
sir_calibration$model_return_prior

valid <- sir_calibration$model_return_prior %>%
	purrr::map(function(x) {
		
		y <- x %>%
			group_by(SampleID) %>%
			select(SampleID, Exp_ID, location_name, treatment_name, treatment_number, year, soc_total, soc_tha_30cm) %>%
			na.omit() %>%
			group_split
		
		yy <- y %>%
			purrr::map(function(data) {
				identifier <- data %>%
					select(SampleID, Exp_ID, location_name, treatment_name, treatment_number) %>%
					unique
				validation_stocks <- validation_calculate_stats(simulated = data[["soc_total"]]/10,
																								 measured = data[["soc_tha_30cm"]]) %>%
					enframe %>%
					pivot_wider(names_from = name) %>%
					rename_with(~ paste0(., "_stocks"))
				data_delta <- data %>%
					mutate(soc_total_delta = soc_total - first(soc_total),
								 soc_tha_30cm_delta = soc_tha_30cm - first(soc_tha_30cm))
				validation_delta <- validation_calculate_stats(simulated = data_delta[["soc_total_delta"]]/10,
																								 measured = data_delta[["soc_tha_30cm_delta"]]) %>%
					enframe %>%
					pivot_wider(names_from = name) %>%
					rename_with(~ paste0(., "_delta"))
				
				return(bind_cols(identifier, validation_stocks, validation_delta))
			}) %>%
			bind_rows
		
		return(yy)}
	) %>%
	bind_rows %>%
	mutate(dist = "Prior Calibration")

```



```{r}
validation_calibration_prior <- validation(site_data = calibration_data,
																					 climate_data = climate_data,
																					 initial_c = initial_c_calibration_spinup,
																					 distribution = sir_calibration$prior,
																					 simulated_column_name = "soc_total",
																					 measured_column_name = "soc_tha_30cm",
																					 model = "ipcct2") %>%
	mutate(dist = "Prior Calibration")
validation_calibration_posterior <- validation(site_data = calibration_data,
																							 climate_data = climate_data,
																							 initial_c = initial_c_calibration_spinup,
																							 distribution = sir_calibration$posterior,
																							 simulated_column_name = "soc_total",
																							 measured_column_name = "soc_tha_30cm",
																							 model = "ipcct2") %>%
	mutate(dist = "Posterior Calibration")

validation_calibration <- bind_rows(validation_calibration_prior,
																		validation_calibration_posterior)
```

## Graphing validation statistics
```{r}
validation_violin_graph <- ggplot(validation_calibration, aes(y=rmse, x=dist)) +
	geom_violin() +
	theme_bw()
ggsave(plot = validation_violin_graph, filename = here(experiment_path,"figures","validation.png"))
```


# Uncertainty Analysis
```{r}
uncertainty_calibration_prior <- montecarlo(site_data = calibration_data,
																						climate_data = climate_data,
																						initial_c = initial_c_calibration_spinup,
																						distribution = sir_calibration$prior)
uncertainty_calibration_posterior <- montecarlo(site_data = calibration_data,
																								climate_data = climate_data,
																								initial_c = initial_c_calibration_spinup,
																								distribution = sir_calibration$posterior)

# Prior
readr::write_csv(uncertainty_calibration_prior$median, file=paste0(experiment_path,"/outputs/uncertainty_median_prior.csv"))
readr::write_csv(uncertainty_calibration_prior$montecarlo, file=paste0(experiment_path,"/outputs/uncertainty_montecarlo_prior.csv"))

# Posterior
readr::write_csv(uncertainty_calibration_posterior$median, file=paste0(experiment_path,"/outputs/uncertainty_median_posterior.csv"))
readr::write_csv(uncertainty_calibration_posterior$montecarlo, file=paste0(experiment_path,"/outputs/uncertainty_montecarlo_posterior.csv"))
```

## Graphing
```{r}
# 1:1 simulated:measured graph
uncertainty_graph <- ggplot(data = NULL, aes(x=soc_total/10, y=soc_tha*(30/(soil_depth_max_cm - soil_depth_min_cm)))) +
	geom_point(data=uncertainty_calibration_prior$montecarlo, shape=4, color = "lightgrey") +
	geom_point(data=uncertainty_calibration_posterior$montecarlo, shape=3, color = "darkgrey") +
	geom_point(data=uncertainty_calibration_posterior$median, color="black") +
	geom_abline(intercept = 0, slope = 1) +
	coord_fixed(ratio = 1, xlim = c(0,400), ylim = c(0,400), expand = TRUE, clip="off") +
	theme_bw()

ggsave(plot = uncertainty_graph, filename = here(experiment_path,"figures","uncertainty.png"))
```

```{r}
ribbon_points_posterior <- uncertainty_calibration_posterior$montecarlo %>%
	group_by(year, treatment_name) %>%
	select(year, treatment_name, soc_total) %>%
	summarise(soc_total_max = max(soc_total),
						soc_total_min = min(soc_total))

ribbon_points_prior <- uncertainty_calibration_prior$montecarlo %>%
	group_by(year, treatment_name) %>%
	select(year, treatment_name, soc_total) %>%
	summarise(soc_total_max = max(soc_total),
						soc_total_min = min(soc_total))

ggplot(data = NULL, aes(x=year)) +
	
	geom_ribbon(data=ribbon_points_prior, aes(ymax=soc_total_max/10, ymin=soc_total_min/10), fill="lightgrey") +
	geom_ribbon(data=ribbon_points_posterior, aes(ymax=soc_total_max/10, ymin=soc_total_min/10), fill="darkgrey") +
	geom_point(data=uncertainty_calibration_posterior$montecarlo, aes(y=soc_tha*(30/(soil_depth_max_cm - soil_depth_min_cm))), color = "black") +
	geom_line(data=uncertainty_calibration_posterior$median, aes(y=soc_total/10), color = "black") +
	facet_wrap(vars(treatment_name)) +
	theme_bw()

```

# Save all code used in this analysis
```{r}
ls()
objects()
```







