---
title: "Run experiment"
author: "Francis Durnin-Vermette"
date: "2023/2/10"
output:
  html_notebook:
    toc: true
    number_sections: true
    df_print: paged
bibliography: bib.bib
editor_options:
  chunk_output_type: inline
---
<!-- Setup chunk: -->
```{r include=FALSE}
library(ggplot2)
library(knitr)
library(dplyr)
library(geojson)
library(tibble)
library(purrr)
library(stringr)
library(here)
library(lubridate)
library(tidyr)
dir.create(tempdir()) #This fixes a bug if the temporary directory is not found
here::i_am("walkthrough_icbm.rmd")
source(here::here("r/icbm_calculate_re.r"))
source(here::here("r/icbm_run.r"))
source(here::here("r/read_climate_data.r"))
source(here::here("r/spinup.r"))
source(here::here("r/bayesian_gsa.r"))
source(here::here("r/bayesian_sir.r"))
source(here::here("r/bayesian_validation.r"))
source(here::here("r/bayesian_uncertainty.r"))
```

Assign experimental parameters:
```{r}
experiment_driver <- "location_name"
experiment_driver_values <- c("Ellerslie", "Harrow")
experiment_model <- "ipcct2"
experiment_time <- Sys.time() %>%
	format("%Y%m%dT%H%M%S")
```

Choose sample sizes for the analysis:
```{r}
calibration_sample_percentage <- 7
gsa_sample_size <- 100
sir_sample_size <- 1000
sir_resample_size <- 100
```

Create folders needed for analysis:
```{r}
experiment_driver_sanitized <- gsub("[^A-Za-z0-9 ]","", experiment_driver)
experiment_driver_values_sanitized <- gsub("[^A-Za-z0-9 ]","",experiment_driver_values)

experiment_name <- paste0(c(experiment_driver_sanitized, experiment_driver_values_sanitized), collapse = "-")

experiment_folder <- paste0(experiment_name, "_", experiment_time, "_", "gsa", gsa_sample_size, "_", "sir", sir_sample_size, "-", sir_resample_size)
experiment_path <- here("results", experiment_model, experiment_folder)

dir.create(here(experiment_path))
dir.create(here(experiment_path, "inputs"))
dir.create(here(experiment_path, "gsa"))
dir.create(here(experiment_path, "sir"))
dir.create(here(experiment_path, "validation"))
dir.create(here(experiment_path, "uncertainty"))
```


# Input data

## Get site data:
```{r}
# Read SLC polygon IDs
polyids_all <- readr::read_csv(here("data","climate_data","Documentation", "SLC_PCPTREGION.csv"))
# Read site data
site_data <- readr::read_csv(here("data","all_experiments_dummy.csv")) %>%
	#site_data <- readr::read_csv(here("data/lte_master_beta_dec02.csv")) %>%
	left_join(polyids_all, by=c("latitude" = "Lat", "longitude" = "Long")) %>%
	mutate(tillage = ifelse(is.na(tillage), "CT", tillage))

# Save to file
readr::write_csv(site_data, file=here(experiment_path,"inputs", "site_data.csv"))
```

## Remove sites where there are no measured values
```{r}
site_data_filtered <- site_data %>%
	group_by(location_name, treatment_name, replication_number) %>%
	filter(!any(is.na(soc_tha_30cm)))

```


## Only select data associated with the driver:
```{r}
site_data_driver <- site_data_filtered %>%
	filter(.data[[experiment_driver]] %in% experiment_driver_values)

# Save to file
site_data_driver %>%
	readr::write_csv(file=here(experiment_path,"inputs","site_data_driver.csv"))
```


## Split the data into calibration and validation sets:
```{r}
site_data_driver_n <- site_data_driver %>%
	group_by(location_name, treatment_name, replication_number) %>%
	n_groups

calibration_sample_indices <- sample.int(site_data_driver_n, ceiling(site_data_driver_n * calibration_sample_percentage/100))
validation_sample_indices <- setdiff(1:site_data_driver_n, calibration_sample_indices)


calibration_data <- site_data_driver %>%
	group_by(location_name, treatment_name, replication_number) %>%
	group_split %>%
	`[`(calibration_sample_indices)

validation_data <- site_data_driver %>%
	group_by(location_name, treatment_name, replication_number) %>%
	group_split %>%
	`[`(validation_sample_indices)

# Save to file
calibration_data %>%
	bind_rows() %>%
	readr::write_csv(file=here(experiment_path,"inputs","site_data_calibration.csv"))
validation_data %>%
	bind_rows() %>%
	readr::write_csv(file=here(experiment_path,"inputs","site_data_validation.csv"))
```

## Get climate data:
```{r}
climate_data <- read_climate_data(climate_data_directory = here("data","climate_data","W9param_TablesCleaned"))

# Save to file
readr::write_csv(climate_data, file=here(experiment_path,"inputs","climate_data.csv"))
```

## Parameter ranges used in sensitivity analysis and bayesian calibration:
```{r}
param_bounds <- read.csv(here("data","ipcct2_parameters_gsa.csv"), stringsAsFactors = FALSE)

# Save to file
readr::write_csv(param_bounds, file=here(experiment_path,"inputs","parameter_ranges.csv"))
```

## Run spinup:
```{r}
calibration_data_spinup <- calibration_data %>%
	map(~slice(., 1:10))

# Get initial C
initial_c_calibration_measured <- calibration_data %>%
	map(~.$initial_soc_tha[1]*10) %>%
	map(~ifelse(is.na(.), 1000, .))

# Run spinup
initial_c_calibration_spinup <- map2(.x = calibration_data_spinup,
																		 .y = initial_c_calibration_measured,
																		 ~spinup(site_data = .x, 
																		 				climate_data = climate_data,
																		 				initial_c = .y,
																		 				model = "ipcct2"))

```

# Sensitivity Analysis
```{r}
# Sobol-Jansen'
sensitivity_calibration_sj <- gsa(site_data = calibration_data,
																	climate_data = climate_data,
																	initial_c = initial_c_calibration_spinup,
																	parameter_bounds = param_bounds,
																	sample_size = gsa_sample_size,
																	method = "soboljansen")
# Save to file
sensitivity_calibration_sj %>%
	mutate(driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 n = gsa_sample_size) %>%
	readr::write_csv(file=here(experiment_path,"gsa","sj.csv"))

# Fast99
sensitivity_calibration_f99 <- gsa(site_data = calibration_data,
																	 climate_data = climate_data,
																	 initial_c = initial_c_calibration_spinup,
																	 parameter_bounds = param_bounds,
																	 sample_size = gsa_sample_size,
																	 method = "fast99")
# Save to file
sensitivity_calibration_f99 %>%
	mutate(driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 n = gsa_sample_size) %>%
	readr::write_csv(file=here(experiment_path,"gsa","f99.csv"))
```

## Graphing:
```{r}
# First-order sensitivity
sensitivity_graph_first <- ggplot(sensitivity_calibration, aes(x = reorder(params, -singsi), y = singsi, ymax=singsi.uci, ymin=singsi.lci)) + 
	xlab("Parameters") +
	ylab("First-Order Sensitivity Index") +
	geom_errorbar(width=0.2, linewidth=1, color="black") +
	geom_bar(stat='identity', fill="grey", alpha=0.70 ) +
	coord_flip() +
	theme_bw()

# Total order sensitivity
sensitivity_graph_total <- ggplot(sensitivity_calibration, aes(x = reorder(params, -singsi), y = totsi, ymax=totsi.uci, ymin=totsi.lci))+ 
	xlab("Parameters") +
	ylab("Total-Order Sensitivity Index") + 
	geom_errorbar(width=0.2, linewidth=1, color="black")+
	geom_bar(stat='identity', fill="grey", alpha=0.70) +
	coord_flip() +
	theme_bw()

# Both
sensitivity_graph_both <- ggplot(sensitivity_calibration, aes(x = reorder(params, -singsi), y = totsi, ymax=totsi.uci, ymin=totsi.lci)) +
	geom_bar(aes(y = singsi), stat='identity', fill="black", alpha=0.7)+
	geom_bar(aes(y = totsi), stat='identity', fill="grey", alpha=0.7)+
	geom_errorbar(width=0.2, linewidth=1)+
	coord_flip() +
	theme_bw()

ggsave(plot = sensitivity_graph_first, filename = here(experiment_path,"gsa","first.png"))
ggsave(plot = sensitivity_graph_total, filename = here(experiment_path,"gsa","total.png"))
ggsave(plot = sensitivity_graph_both, filename = here(experiment_path,"gsa","both.png"))
```

# Sampling Importance Resampling (SIR)
```{r}
sir_calibration <- sir(site_data = calibration_data,
											 climate_data = climate_data,
											 initial_c = initial_c_calibration_spinup,
											 parameter_bounds = param_bounds,
											 sample_size = sir_sample_size,
											 resample_size = sir_resample_size)

readr::write_csv(sir_calibration$prior, file=here(experiment_path, "sir", "prior.csv"))
readr::write_csv(sir_calibration$posterior, file=here(experiment_path, "sir", "posterior.csv"))

# Find maximum a posteriori (MAP)
get_map <- function(x) {
	dens <- x %>%
		density
	return(dens$x[which.max(dens$y)])
}

sir_table <- sir_calibration$posterior %>%
	pivot_longer(cols = !SampleID) %>%
	group_by(parameter = name) %>%
	summarise(`2.5%` = quantile(value, probs = 0.025),
						`25%` = quantile(value, probs = 0.25),
						median = quantile(value, probs = 0.50),
						`75%` = quantile(value, probs = 0.75),
						`97.5%` = quantile(value, probs = 0.975),
						map = get_map(value))

sir_table %>%
	mutate(driver = experiment_driver,
				 value = paste(experiment_driver_values, collapse=","),
				 n_sample = sir_sample_size,
				 n_resample = sir_resample_size) %>%
	readr::write_csv(file= here(experiment_path,"sir","summary.csv"))
```

## Graphing:
```{r}
prior <- sir_calibration$prior %>%
	pivot_longer(cols = !`SampleID`)
posterior <- sir_calibration$posterior %>%
	pivot_longer(cols = !`SampleID`)

sir_graph <- ggplot() +
	geom_density(data = posterior, aes(value), col = NA, fill = "red", alpha = 0.2) +
	geom_density(data = prior, aes(value), col = NA, fill = "blue", alpha = 0.2) +
	facet_wrap(~name, scales = "free", ncol = 3) +
	theme_bw() +
	theme(panel.border = element_rect(colour = "black", fill = NA), panel.grid.major = element_blank(),
				axis.title.x=element_blank(),panel.grid.minor = element_blank()) +
	scale_y_continuous(expand = c(0,0)) + scale_x_continuous(expand = c(0,0))

ggsave(plot = sir_graph, filename = here(experiment_path,"sir","sir.png"))
```

# Validation

## Graphing results
```{r}
stocks_calibration_prior <- sir_calibration$model_return_prior %>%
	purrr::map(~validation(.)$stocks) %>%
	bind_rows

stocks_calibration_prior_soc_graph <- ggplot(stocks_calibration_prior, aes(x = year)) +
	geom_point(aes(y = soc_total)) +
	geom_point(aes(y = soc_tha_30cm), color="red") +
	facet_wrap(vars(treatment_name)) +
	ylab("SOC (t/ha)") +
	theme_bw()
ggsave(plot = stocks_calibration_prior_soc_graph, filename = here(experiment_path,"validation","stocks_calibration_prior_soc.png"))

stocks_calibration_prior_soc_delta_cumulative_graph <- ggplot(stocks_calibration_prior, aes(x = year)) +
	geom_point(aes(y = soc_total_delta_cumulative)) +
	geom_point(aes(y = soc_tha_30cm_delta_cumulative), color="red") +
	facet_wrap(vars(treatment_name)) +
	ylab("ΔSOC (t/ha)") +
	theme_bw()
ggsave(plot = stocks_calibration_prior_soc_delta_cumulative_graph, filename = here(experiment_path,"validation","stocks_calibration_prior_soc_delta_cumulative.png"))


stocks_calibration_prior_soc_delta_progressive_graph <- ggplot(stocks_calibration_prior, aes(x = year)) +
	geom_point(aes(y = soc_total_delta_progressive)) +
	geom_point(aes(y = soc_tha_30cm_delta_progressive), color="red") +
	facet_wrap(vars(treatment_name)) +
	ylab("Year-by-year ΔSOC (t/ha)") +
	theme_bw()
ggsave(plot = stocks_calibration_prior_soc_delta_progressive_graph, filename = here(experiment_path,"validation","stocks_calibration_prior_soc_delta_progressive.png"))

```

## Calculating Validation statistics
```{r}
validation_calibration_prior <- sir_calibration$model_return_prior %>%
	purrr::map(~validation(.)$validation) %>%
	bind_rows %>%
	mutate(dist = "Prior Calibration")

validation_calibration_posterior <- sir_calibration$model_return_posterior %>%
	purrr::map(~validation(.)$validation) %>%
	bind_rows %>%
	mutate(dist = "Posterior Calibration")

# Save to file
readr::write_csv(validation_calibration_prior, file = here(experiment_path,"validation","validation_calibration_prior.csv"))
readr::write_csv(validation_calibration_posterior, file = here(experiment_path,"validation","validation_calibration_posterior.csv"))
```

## Graphing validation statistics
```{r}
validation_calibration <- bind_rows(validation_calibration_prior, validation_calibration_posterior)

validation_violin_graph <- ggplot(validation_calibration, aes(y=rmse_stocks, x=dist)) +
	geom_violin() +
	theme_bw()
ggsave(plot = validation_violin_graph, filename = here(experiment_path,"validation","validation.png"))
```


# Uncertainty Analysis
```{r}
uncertainty_calibration_prior <- montecarlo(site_data = calibration_data,
																						climate_data = climate_data,
																						initial_c = initial_c_calibration_spinup,
																						distribution = sir_calibration$prior)
uncertainty_calibration_posterior <- montecarlo(site_data = calibration_data,
																								climate_data = climate_data,
																								initial_c = initial_c_calibration_spinup,
																								distribution = sir_calibration$posterior)

# Write to file
# Prior
readr::write_csv(uncertainty_calibration_prior$median, file=here(experiment_path,"uncertainty", "median_prior.csv"))
readr::write_csv(uncertainty_calibration_prior$montecarlo, file=here(experiment_path,"uncertainty", "montecarlo_prior.csv"))
# Posterior
readr::write_csv(uncertainty_calibration_posterior$median, file=here(experiment_path,"uncertainty", "median_posterior.csv"))
readr::write_csv(uncertainty_calibration_posterior$montecarlo, file=here(experiment_path,"uncertainty", "montecarlo_posterior.csv"))
```

## Graphing
```{r}
# 1:1 simulated:measured graph
uncertainty_graph <- ggplot(data = NULL, aes(x=soc_total/10, y=soc_tha*(30/(soil_depth_max_cm - soil_depth_min_cm)))) +
	geom_point(data=uncertainty_calibration_prior$montecarlo, shape=4, color = "lightgrey") +
	geom_point(data=uncertainty_calibration_posterior$montecarlo, shape=3, color = "darkgrey") +
	geom_point(data=uncertainty_calibration_posterior$median, color="black") +
	geom_abline(intercept = 0, slope = 1) +
	coord_fixed(ratio = 1, xlim = c(0,400), ylim = c(0,400), expand = TRUE, clip="off") +
	theme_bw()

ggsave(plot = uncertainty_graph, filename = here(experiment_path,"uncertainty","uncertainty.png"))
```

```{r}
ribbon_points_posterior <- uncertainty_calibration_posterior$montecarlo %>%
	group_by(year, treatment_name) %>%
	select(year, treatment_name, soc_total) %>%
	summarise(soc_total_max = max(soc_total),
						soc_total_min = min(soc_total))

ribbon_points_prior <- uncertainty_calibration_prior$montecarlo %>%
	group_by(year, treatment_name) %>%
	select(year, treatment_name, soc_total) %>%
	summarise(soc_total_max = max(soc_total),
						soc_total_min = min(soc_total))

ggplot(data = NULL, aes(x=year)) +
	
	geom_ribbon(data=ribbon_points_prior, aes(ymax=soc_total_max/10, ymin=soc_total_min/10), fill="lightgrey") +
	geom_ribbon(data=ribbon_points_posterior, aes(ymax=soc_total_max/10, ymin=soc_total_min/10), fill="darkgrey") +
	geom_point(data=uncertainty_calibration_posterior$montecarlo, aes(y=soc_tha*(30/(soil_depth_max_cm - soil_depth_min_cm))), color = "black") +
	geom_line(data=uncertainty_calibration_posterior$median, aes(y=soc_total/10), color = "black") +
	facet_wrap(vars(treatment_name)) +
	theme_bw()

```






